Cartogram Mapping and its Application to Cancer Data Visualisation

### 2019-09-24 review_article.Rmd
animation figure

An intentional goal when creating the 1966 Census population cartogram for Canada was to maintain contiguity, while attempting to keep the actual shape of places. The end result was a 'very accurate isodemo-graphic map of Canada'. This intentional design goal coincided with the rising interest in urban geography and presentation of social statistics.

To be able to recognise the significant changes, a reader will usually have to know the initial geography to find the differences in the new cartogram layout [@NAC]. Tobler's Conformal mapping means to preserve angles locally so that the shapes of small areas on a traditional map and a cartogram would be similar.
@CBATCC presents this issue as conflicting tasks or aims, to adjust region sizes and retain region shapes. 
@MDAC define the term 'mosaic cartograms' for hexagonal tile displays, where the number of tiles for each area can be used to communicate the statistic of regions. 
The complexity of the boundaries can be adjusted in the resulting display, as the size of the tiles used allows a trade-off be made between boundary complexity and simplicity.

<!--
Tobler's method and the many implementations that 'elaborated' on it are derived from 'numerical approximations to a pair of equations'[@ACTUC]. They all operate through incremental adjustments, and can produce wildly different outcomes from small changes in the inputs.

Computer generated map examples:  @HEDP
(Hopps et al. 1968; Armstrong 1972). There
has followed a flood of disease atlases, mainly concentrating on the modem problems of cancer and degenerative diseases from countries as scattered as the United States (Burbank 1971; Mason et al. 1975, 1976; Pickle et al. 1987), the Soviet Union (Levin 1980), Japan (Shigematsu 1977), the Federal Republic of Germany
-->

@@

As the trade-off regarding boundaries approaches simplicity, the distortion of region shapes on the contiguous cartogram presents an additional hurdle to visual recognition and this hurdle is not only eliminated on the non-contiguous cartogram but is replaced by the meaningful empty-space property [@NAC, @ECGC].
The shapes are valuable for recognition and allows users to orient themselves on the display. Map creators can efficiently communicate with this kind of map by keeping the outlines or particular elements of the original in the new shape [@NISCC].
The scale of the areas does not impact on the shape recognition. However it may impact on the visibility of all areas if small areas expand beyond their boundaries.

@@

He acknowledged the sophistication of contiguous cartograms but critiqued their 'very complex shapes,' he answers this with his implementation of maps created using 'the simplest of all shapes'. 
Circular cartograms use the same simple shapes for every region represented, and resizes the shapes according to the statistic represented or the population for a base map. This familiar shape may be more effective for understanding the spatial distribution than contiguous cartograms, as the 'nonsense' shapes used have 'no meaning' after distortions are applied @NISCC.
To produce a compelling map, a gravity model is applied to avoid overlaps, and keep spatial relationships with neighbouring areas over many iterations. This implementation can work for up to 'one hundred thousand' areas.

@@

A similar method, where each geographic area is represented using a square, tessellated to create a square grid. This method has been used by FiveThirtyEight, Bloomberg Business, The Guardian, The Washington Post, The New York Times and NPR.
Each area is represented by a square of the same dimensions.
Figure \ref{fig:ggcartograms} d shows a hexagon tile map representing each state using a single square.

@@

A simple tile map presents the areas in a tessellated grid display, where geographic neighbours are found next to each other, with some necessary displacement employed for regions with more than four neighbours.
These tiles may be labelled or coloured to represent a value.
Tile maps may be difficult to create, they are best created manually, with additional time and care required as the amount of geographic areas to include increases.

@@

>>>>

An intentional goal when creating the 1966 Census population cartogram for Canada was to maintain contiguity, while attempting to keep the actual shape of places. The end result was a 'very accurate isodemo-graphic map of Canada'. This intentional design goal coincided with the rising interest in urban geography and presentation of social statistics. Figure \ref{fig:ggcartograms} a shows a population contiguous cartogram of the United States. All states are visible and the shape of the United States overall is still recognisable.
In contrast, Figure \ref{fig:auscartograms} a shows an Australian contiguous cartogram also based on population. The south east is completely distorted, areas with low popuation are still large very the map, as their initial area was used to determine the area allocated in the contiguous cartogram.

To be able to recognise the significant changes, a reader will usually have to know the initial geography to find the differences in the new cartogram layout [@NAC]. Tobler's Conformal mapping means to preserve angles locally so that the shapes of small areas on a traditional map and a cartogram would be similar.
@CBATCC presents this issue as conflicting tasks or aims, to adjust region sizes and retain region shapes.

@@

As the trade-off regarding boundaries approaches simplicity, the distortion of region shapes on the contiguous cartogram presents an additional hurdle to visual recognition and this hurdle is not only eliminated on the non-contiguous cartogram but is replaced by the meaningful empty-space property [@NAC, @ECGC].
The shapes are valuable for recognition and allows users to orient themselves on the display. Map creators can efficiently communicate with this kind of map by keeping the outlines or particular elements of the original in the new shape [@NISCC].
The scale of the areas does not impact on the shape recognition.
However it may impact on the visibility of all areas if small areas expand beyond their boundaries. 
Figure \ref{fig:ggcartograms} b shows small overlaps in the north east region of the United States. These overlaps are also shown in the non-contiguous cartogram of Australia in Figure \ref{fig:auscartograms} b.

@@

He acknowledged the sophistication of contiguous cartograms but critiqued their 'very complex shapes,' he answers this with his implementation of maps created using 'the simplest of all shapes'. 
Circular cartograms use the same simple shapes for every region represented, and resizes the shapes according to the statistic represented or the population for a base map. This familiar shape may be more effective for understanding the spatial distribution than contiguous cartograms, as the 'nonsense' shapes used have 'no meaning' after distortions are applied @NISCC.
To produce a compelling map, a gravity model is applied to avoid overlaps, and keep spatial relationships with neighbouring areas over many iterations. This implementation can work for up to 'one hundred thousand' areas.
Figure \ref{fig:ggcartograms} c shows the displacement of the north east region of the United States in the Dorling cartogram, the rest of the country is still recognisable.
This displacement is shown to an extreme in \ref{fig:auscartograms} c the Dorling cartogram of Australia.

@@

A similar method, where each geographic area is represented using a square, tessellated to create a square grid. This method has been used by FiveThirtyEight, Bloomberg Business, The Guardian, The Washington Post, The New York Times and NPR.
Each area is represented by a square of the same dimensions.
Figure \ref{fig:ggcartograms} d shows a hexagon tile map representing each state using a single hexagon.

@@

A simple tile map presents the areas in a tessellated grid display, where geographic neighbours are found next to each other, with some necessary displacement employed for regions with more than four neighbours.
These tiles may be labelled or coloured to represent a value.
Tile maps may be difficult to create, they are best created manually, with additional time and care required as the amount of geographic areas to include increases.
@MDAC define the term 'mosaic cartograms' for hexagonal tile displays, where the number of tiles for each area can be used to communicate the statistic of regions. 
The complexity of the boundaries can be adjusted in the resulting display, as the size of the tiles used allows a trade-off be made between boundary complexity and simplicity.

@@

```{r animating, fig.cap = "Animations in publicly available choropleth cancer maps. [a. Map of Cancer Mortality Rates in Spain](http://www.geeitema.org/AtlasET/atlas.jsp?causa=e27MEPOC)(Controls for indicator, sex, age groups, continents, and cancer types) , [b. Breast Cancer Mortality in Canada](http://www.ehatlas.ca/light-pollution/maps/breast-cancer-mortality) (Controls for indicator, age, year, and geographical areas)", fig.width = 12, fig.height = 9}
a1 <- png::readPNG("cancer_map_images/animation/23.animation.png")
a2 <- png::readPNG("cancer_map_images/animation/2.animation.png")

ggdraw() +
  draw_plot(rasterGrob(a1), 0, 0, 0.3, 1) +
  draw_plot(rasterGrob(a2), 0.3, 0, 0.7, 1) +
  draw_plot_label(c("I", "II"), 
    c(0, .3), c(1, 1), size = 20)
```
